---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
*Loading and processing data to a suitable format for analysis*
```{r}
unzip('activity.zip')
data <- read.csv('activity.csv', colClasses = c('integer', 'Date', 'factor'))
data$month <- as.numeric(format(data$date, '%m'))
clean <- na.omit(data)
rownames(clean) <- 1:nrow(clean)
```


## What is mean total number of steps taken per day?
*Total number of steps taken per day*
```{r}
total_steps <- aggregate(clean$steps, list(Date = clean$date), FUN  = 'sum')$x
total_steps
```
*Histogram of total steps taken per day*
```{r}
library(ggplot2)
ggplot(clean, aes(date, steps)) + geom_bar(stat = 'identity' ) + labs(title = 'histogram of total number of steps taken per day', x = 'Date', y = 'Number of steps')
```
*Mean of total steps*
```{r}
 ogmean <- mean(total_steps)
ogmean
```
 *Median of total steps*
```{r}
ogmedian <- median(total_steps)
ogmedian
```
## What is the average daily activity pattern?
*Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days
```{r}
Average_tests <- aggregate(clean$steps, list(interval = as.numeric(as.character(clean$interval))), FUN = 'mean')
names(Average_tests)[2] <- 'meanofsteps'
ggplot(Average_tests, aes(interval, meanofsteps)) + geom_line(color = 'black') + labs(title = "Time Series Plot of the 5-minute Interval", x = "5-minute intervals", y = "Average Number of Steps Taken")
```
```
## Imputing missing values
*Total number of missing values in the data set*
```{r}
sum(is.na(data))
```
*Replacing all missing values with the median and creating a new data set containing these new values*
```{r}
newData <- data 
for (i in 1:nrow(newData)) {
  if (is.na(newData$steps[i])) {
    newData$steps[i] <- Average_tests[which(newData$interval[i] == Average_tests$interval), ]$meanofsteps
  }
}

```
*Creating a histogram with the new data, which has the missing values filled in *
```{r}
ggplot(newData, aes(date, steps)) + geom_bar(stat = "identity") + labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")
```
*Difference in the values of mean and median as compared to previous data*
```{r}

newTotalSteps <- aggregate(newData$steps,list(Date = newData$date), FUN = "sum")$x

newmean <- mean(newTotalSteps)
newmean
newmedian <- median(newTotalSteps)
newmedian

newmean - ogmean
newmedian - ogmedian
```
## Are there differences in activity patterns between weekdays and weekends?
*Creating an new factor variable indication whether a day is a weekend or weekday
```{r}
newData$weekdays <- factor(format(newData$date, "%A"))
levels(newData$weekdays)
levels(newData$weekdays) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))
levels(newData$weekdays)
table(newData$weekdays)
```
*Creating a plot of the 5-Minute interval and average number of steps taken, avergaed across weekdays or weekend days*
```{r}
avgSteps <- aggregate(newData$steps,list(interval = as.numeric(as.character(newData$interval)) , weekdays = newData$weekdays), FUN = "mean")
names(avgSteps)[3] <- "meanOfSteps"
library(lattice)
xyplot(avgSteps$meanOfSteps ~ avgSteps$interval | avgSteps$weekdays,layout = c(1, 2), type = "l", xlab = "Interval", ylab = "Number of steps")
```

